### 2024-11-21 paper/dsl.tex
Add al and dsl style WIP: write dsl

>>>>

% DSL high-level explanation
SpecTec offers a domain-specific language (DSL) tailored for specification
purposes.
It is a declarative language, providing a compact and user-friendly syntax.
It consists of DSL definitions: type, grammar, relation, and functions.
Type definitions are used to define abstract syntax, while grammar definitions
are used to define concrete syntax.
The below code represents the abstract syntax and concrete syntax for
\texttt{numtype}.
\begin{lstlisting}[style=dsl]
syntax numtype = I32 | I64 | F32 | F64 | V128
grammar Bnumtype : numtype =
  | 0x7F => I32
  | 0x7E => I64
  | 0x7D => F32
  | 0x7C => F64
\end{lstlisting}
Relation definitions are used to define semantics, while function definitions
are used auxiliary definitions.

It is type checked to prevent meta-level errors such as notation misuses and
dimension mismatches. The code below is the specification for \texttt{testop}
instruction written in the DSL, which is designed to be similar to the formal
notation when describing the semantics with compact and user-friendly
notation.
\begin{lstlisting}[style=dsl]
  rule Step_pure/testop:
  (CONST nt c_1) (TESTOP nt testop)  @$\tildecode$@>  (CONST I32 c)
  -- if c = @$\dollarcode$@testop_(nt, testop, c_1)
\end{lstlisting}

% DSL
The DSL consists of DSL definitions.
There are mainly 4 kinds of DSL definitions: type, grammar, relation, and
functions.
Type definitions are used to define abstract syntax, while grammar definitions
are used to define concrete syntax.
Relation definitions are used to define semantics; the code above is also
described in relation definition.
Function definitions can be utilized for other definitions such as
\texttt{\$testop\_} in the code.
In addition, they can be also used to describe semantics.
Especially, SpecTec uses function definitions to describe module instantiation
and function invocation.

### 2024-11-21 paper/overview.tex
Write background
Add al and dsl style WIP: write dsl

% Overview explanation
In this chapter, we explain the overall approach to generating multiple
artifacts from the specification and ensuring the correctness of the
specification.
\Cref{fig:overview} illustrates an overview of our approach.
% SpecTec: specification mechanization tool
SpecTec is a tool for \textit{mechanizing} WebAssembly specification
~\cite{spectec}.
Specification mechanization is a technique that treats a specification as data
that can be manipulated by a machine to automatically generate parser and
interpreter ~\cite{jiset}, to generate tests and perform differential tests
~\cite{jest}, to find meta-level errors in the specification ~\cite{jstar}, and
to perform meta-level static analysis for the static analysis of the defined
language ~\cite{jsaver}.
This technique is especially crucial in growing languages such as JavaScript
and WebAssembly.

% DSL
SpecTec provides a domain-specific language (DSL) for writing specifications.
It is a declarative language designed to resemble the formal notation while
offering a compact and user-friendly syntax.
The code below shows the specification for the \texttt{testop} instruction
written in the DSL:
\begin{lstlisting}[escapechar=\%]
  rule Step_pure/testop:
  (CONST nt c_1) (TESTOP nt testop)  %$\tildecode$%>  (CONST I32 c)
  -- if c = %$\dollarcode$%testop_(nt, testop, c_1)
\end{lstlisting}
The DSL is type checked to prevent meta-level errors such as notation misuses and
dimension mismatches.
Using the DSL, SpecTec generates LaTeX code and mechanized definitions for theorem provers.

% AL
Additionally, SpecTec introduces another imperative language, \textit{AL}, for
the prose notation.
In AL, semantics are expressed algorithmically, in the form of step-by-step
instructions that take inputs and perform operations, mirroring the prose notation.
SpecTec translates declarative definitions in the DSL into algorithmic
definitions in AL.
The AL code below is the specification for the \texttt{testop} instruction
generated from the DSL:
\begin{lstlisting}[escapechar=\%]
Step_pure/testop nt testop {
 Assert (top_value(nt))
 Pop (numtype_0.CONST c_1)
 Let c = %$\dollarcode$%testop_(nt, testop, c_1)
 Push (I32.CONST c)
}
\end{lstlisting}
Using AL, SpecTec generates reStructuredText code.
\cref{fig:spectec-testop} shows the generated document for \texttt{testop}
instruction.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.

@@

\red{(Explanation for each chapter for SpecTec)}

>>>>

% SpecTec: specification mechanization tool
SpecTec is a tool for \textit{mechanizing} WebAssembly specification
~\cite{spectec}.
Specification mechanization is a technique that treats a specification as data
that can be manipulated by a machine to automatically generate parser and
interpreter ~\cite{jiset}, to generate tests and perform differential tests
~\cite{jest}, to find meta-level errors in the specification ~\cite{jstar}, and
to perform meta-level static analysis for the static analysis of the defined
language ~\cite{jsaver}.
% Overview explanation
In this chapter, we explain the overall approach to generating multiple
artifacts from the specification and ensuring the correctness of the
specification.
\Cref{fig:overview} illustrates an overview of our approach.

% DSL
SpecTec offers a domain specific language (DSL) designed for writing
Wasm specifications.
It is a declarative language, providing a compact and user-friendly syntax.
This DSL can represent Wasm semantics in a manner closely resembling the formal
notation as shown in the following code for specification of testop
instruction.
The code below shows the specification for the \texttt{testop} instruction
written in the DSL:
\begin{lstlisting}[style=dsl]
  rule Step_pure/testop:
  (CONST nt c_1) (TESTOP nt testop)  @$\tildecode$@>  (CONST I32 c)
  -- if c = @$\dollarcode$@testop_(nt, testop, c_1)
\end{lstlisting}
The DSL is type checked to prevent meta-level errors such as notation misuses and
dimension mismatches.
Using the DSL, SpecTec generates LaTeX code and mechanized definitions for theorem provers.

% AL
Additionally, SpecTec introduces another imperative language, \textit{AL}, for
the prose notation.
In AL, semantics are expressed algorithmically, in the form of step-by-step
instructions that take inputs and perform operations, mirroring the prose notation.
SpecTec translates declarative definitions in the DSL into algorithmic
definitions in AL.
The AL code below is the specification for the \texttt{testop} instruction
generated from the DSL:
\begin{lstlisting}[style=al]
@\textbf{Step\_pure/testop}@ nt testop {
    Assert (top_value(nt))
    Pop (numtype_0.CONST c_1)
    Let c = @$\dollarcode$@testop_(nt, testop, c_1)
    Push (I32.CONST c)
}
\end{lstlisting}
Using AL, SpecTec generates reStructuredText code.
\cref{fig:spectec-testop} shows the generated document for \texttt{testop}
instruction.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.

@@
